---
title: "TidyTemplate"
date: 2022-03-16
output: html_document
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}


library(tidyverse)
library(tidytuesdayR)
library(lubridate)
# library(ggfortify)
library(ggtext)

base_path = here::here("2022","11")


```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2022-03-15")


```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

## sep the two datasets out

cran <- tt$cran
bioc <- tt$bioc

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

The glaring issue with the dataset is in the `cran` table. The dates are clearly in different formats. Let's try to get those looking nice. 

```{r Wrangle}

## using @geokaramanis method for cleanign dates
cran %>% 
    mutate(version_dot = str_replace_all(version, "-", ".")) %>% 
  filter(date != 0) %>% 
  mutate(
    date_ymd_utc = parse_date(date, format = c("%Y-%m-%d %H:%M:%S UTC")),
    date_ymd_cdt = parse_date(date, format = c("%Y-%m-%d %H:%M:%OS CDT")),
    date_dmy = parse_date(date, format = c("%a %b %d %H:%M:%S %Y")),
    date_ymd = parse_date(date, format = c("%Y-%m-%d"))
  ) %>% 
  pivot_longer(starts_with("date_"), values_to = "date_short") %>%
  # remove one package ("PIN") that has date in 1987:
  filter(date_short > "2000-12-31") %>% 
  select(-name) %>% 
  group_by(package) %>% 
  arrange(date_short) %>% 
  mutate(package_idx = cur_group_id()) %>% 
  arrange(package_idx) %>% 
  mutate(
    version_idx = row_number(),
    days_since_start = as.numeric(date_short - date_short[1]),
    days_since_last_update = as.numeric(date_short - lag(date_short)),
    days_to_today = as.numeric(Sys.Date()- date_short),
  ) %>% 
  ungroup() ->
  cran_clean

cran_clean %>% 
  group_by(package) %>% 
  ## get rid of packages updated in the last year
  filter(!any(days_to_today < 365*2)) %>% 
  replace_na(list(days_since_last_update = 0)) %>% 
  mutate(will_there_be_update = 1:n() != n()) %>% 
  ungroup() %>% 
  mutate(time_bin = cut(days_since_start, breaks = seq(0, 365*2, 30),right = F)) %>% 
  group_by(time_bin) %>% 
  mutate(last_update = mean(!will_there_be_update), n = n(),
            min_time = min(days_since_start)) ->
  surv_data



## grab two examples

surv_data %>% 
  filter(min_time == 0, !will_there_be_update, package == "Bhat") ->
  zero_ex

surv_data %>% 
  filter(between(min_time, 300, 500), !will_there_be_update, package == "ggplot")  ->
  mid_ex

```


# Visualize

My thought is to look for the "Lindy Effect" in the version updates of packages. From Wikipedia:  

> The Lindy effect (also known as Lindy's Law[1]) is a theorized phenomenon by which the future life expectancy of some non-perishable things, like a technology or an idea, is proportional to their current age.  



```{r Visualize}

# Annotations
plot_titles <- list(
  title = "Getting started is the hardest part",
  subtitle = "Each dot is 50 package updates based on the package's lifespan at the time.<br>While packages slowly stop being updated, the most likely **last** update is the **first** update.",
  caption = "
  **Source:** Robert Flight (@rmflight) | **Visualization:** Doug Liebe (@dougliebe)"
)

jitter <- position_jitter( height = 0.01, width = 10)
jitter2 <- position_jitter( height = 0.01, width = 10)
surv_data %>% 
  group_by(time_bin) %>% 
  mutate(n = round(n()/50)) %>% 
  sample_n(n[1]) %>% 
  ggplot()+
  geom_jitter(aes(min_time, last_update, color = n), alpha = 0.5, height = 0.01, width = 10)+
  geom_jitter(data = zero_ex, aes(min_time, last_update), color = "black", position = jitter, size =3)+
  geom_jitter(data = mid_ex, aes(min_time, last_update), color = "red", position = jitter2, size =3)+
  scale_color_viridis_c()+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  labs(
    x = "Days Since First Version Published",
    y = "Will this be the last update?",
    title = plot_titles$title,
    subtitle = plot_titles$subtitle,
    caption = plot_titles$caption
  )+
  theme_minimal()+
  theme(
    legend.position = 'none',
    plot.background = element_rect(fill = 'white', color = 'white'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 32, margin = margin( b = 1)),
    plot.title.position = "plot",
    plot.subtitle = element_markdown(
      size = 18, margin = margin(t = 4, b = 8),lineheight = 0
    ),
    plot.caption = element_markdown(
      margin = margin(t = 8), size = 16, color = "grey30"),
    axis.text = element_text(size =16),
    axis.title = element_text(size = 18),
    text = element_text(color = "grey4", lineheight = 1.25),
    
    plot.margin = margin(t = 8, b = 4, l = 8, r = 18)
  )+
  ## annotations
  annotate("label", x = 20, y = 0.243,
           label = "{Bhat} published on 2006-02-08, no updates", hjust = 0, size = 4) +
  annotate("label", x = 450, y = 0.154, color = 'red',
           label = "{ggplot} last update was 2007-07-09 after 460d", hjust = 0, size = 4) +
  annotate("label", x = 700, y = 0.23, 
           label = "4,437 packages have been updated for 2+ years", hjust = 'right', size = 4) 
  
  

  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(here::here(base_path, "cran_updates.png"), width = 5, height = 3)

```
